### 2020-08-10 text/ecography/round_1/review/reply_to_reviewers.Rmd
migrate gitlab repo to github

---
output:
  bookdown::pdf_document2:
    keep_tex: no
    latex_engine: xelatex
    # float is needed for floating figures
    extra_dependencies: ["fontspec", "sectsty", "float"]
    toc: false
    fig_caption: yes
    number_sections: true
header-includes:
  # adjust figure captions
  - \usepackage{caption}
  - \captionsetup{font={footnotesize,bf,singlespacing}}

title: "Point-by-point reply to reviewers' comments"
bibliography: ../../round_2/transect_refs.bib
csl: ../../round_2/ecography.csl
# biblio-style: "natbib"
link-citations: true

---

```{r setup, include=FALSE}
library("knitr")
library("kableExtra")
library("tidyverse")
library("here")
# make sure that figures are allowed to fload (hold the assigned position and are not always put on top of a page)
knitr::opts_chunk$set(fig.pos = 'H')
```

# Reply to subject editor Bethany Bradley

Dear Dr Bradley,

thank you very much for handling our manuscript.
We are pleased to resubmit a revised version of our manuscript now entitled  *Monitoring and predictive mapping of floristic biodiversity along a climatic gradient in ENSO's terrestrial core region, NW Peru*.
We are also very thankful for the constructive comments of both reviewers which helped to improve substantially our manuscript.
Please find below a point-by-point reply to all concerns raised by the reviewers.
In accordance with your and the reviewers' suggestions, we have:

- formulated clear hypotheses and adjusted accordingly the framing of the manuscript.
- explained our methods in much more detail.
To make our study truly reproducible, we will publish our study upon acceptance as a research compendium.
This means that we will make our code and text publicly available via [github](https://github.com).
The data will be published on DRYAD or Zenodo.
And a docker container will make available the software used during the analyses thereby ensuring that our code will run on any operating system and regardless of software updates.
- embedded our study even better in the ongoing scientific debate while focussing on the key points of our study.
- highlighted more prominently the novelty of our study in the introduction and discussion section.
- written out El Niño/La Niña in full rather than abbreviating.

Therefore, we hope that our manuscript is now suitable for publication.

Yours sincerely,

Dr Jannes Muenchow

# Reply to reviewer #1

1. **The authors assessed the effect of water availability on species diversity across a gradient in the Northwest Peru in different years in which the effect of climatic events (ENSO) have influence on plant diversity. They used 50 (0.09ha) plots distributed in different water availability conditions along a transect and measured plant identity and plant coverage of vascular plants in each one. In addition, they used an irrigation-experiment to assess under controlled conditions the effects of water availability and their interaction with nutrient enrichment on plant diversity accumulation. This integrative approach which combines macroecological analyses and experimental tests of the potential drivers is very interesting and rare in ecological studies. However, the methods and results are not clear. Particularly the description of the plot data, the modeling approaches, and the display of results. Moreover, there is a mismatch in the response variable between the observational analyses and the experiments. Finally, in the introduction, the hypotheses do not have clear predictions, and the discussion focusses on topics that are not presented as the focus of the manuscript in the introduction. In addition, the text should be reviewed to improve the quality of the writing, there are some sentences that are not clear. In my opinion, this study has a lot of potential, but these issues need to be addressed.**

First of all, thank you very much for your detailed feedback which helped to improve substantially the manuscript.
In accordance with your suggestions, we have rewritten the methods and results sections.
Additionally, we have described in much more detail the plot data and the modeling approaches.
We also show that in fact there is no mismatch between the response variables (please refer to the reply to comment #12).
We are sorry that this was not clear from the beginning.
We also have rewritten the research questions as hypotheses with clear predictions.
Additionally, we absolutely agree with you that parts of the discussion in the first draft of the manuscript were out of scope.
Thanks for pointing this out, we have deleted or largely rewritten the corresponding sections and the Conclusions.
Finally, we have improved the writing.

2. **Line 48: The introduction should include a bit about the expectation of species diversity response to climate conditions. Also, the term “ecosystem diversity” should be clarified.**

We have rewritten our research questions as follows:

>>> Specifically, we investigated the:
>>>
(1) change in the different levels of biodiversity, namely alpha, beta, gamma and ecosystem diversity. 
The latter was represented by ordination scores and additionally visualized by means of predictive mapping.
Overall, we expected a boost in biodiversity as soon as water was no longer a limiting factor.
Regarding the spatial distribution of the floristic composition we expected a more pronounced development of vegetation assemblages along the climatic gradient with wetter conditions.
>>>
(2) effects of soil characteristics and topography in shaping floristic composition and vegetation cover (biomass production). 
We expected that the influence of topographic and edaphic variables might gain in influence in wetter years.
>>>
(3) We complemented the plot survey at the landscape scale with an irrigation-fertilization experiment, thereby considering the socio-economic perspective of the ENSO variability. 
Specifically, we quantified the effect of the water-nutrient interaction on biomass development under different rainfall scenarios.
We expected that the water input of the rainfall scenario representing a Super-Niño event in combination with the fertilizer would be most conducive to biomass and biodiversity development.

Additionally, we now specifically say that we represented ecosystem diversity by ordination scores.

3. **Line 50: The introduction is focused on floristic composition and the second question also includes information on biomass production. Here, there are many studies that showed the effect of ENSO and extreme climatic events in aboveground biomass production (e.g. Phillips et al. 2009). Based on these previous studies, the authors could include a bit more of information on the response of biomass production and their expectations for the study system.**

Regarding our expectations, please refer to the reply to the previous comment.
Thanks for pointing us to relevant literature.
We have added following sentence to the introduction:

>>>In any case, ENSO certainly drives a large amount of the interannual variability of woody plant growth across the tropics [@rifai_enso_2018] though the specific response to El Niño may vary from region to region [@phillips_changes_2009].

<!--
- [@phillips_changes_2009] -> mixed response of Amazon forest to EN induced droughts.
However, climate change affects Amazon moist forests, there is a risk that they might be converted from carbon sinks to sources.
- cite also Salazar et al. (accepted) since they also show that there is a risk that climate change might convert TDF from sink to sources.
- [@rifai_enso_2018]: ENSO Drives interannual variation of forest woody growth across the tropics 
Overall, there is high correlation (r = -0.75) between the annual anomaly of tropical forest woody growth and the annual mean of the El Niño 3.4 index, driven mainly by strong correlations with anomalies of soil water deficit, vapour pressure deficit and shortwave radiation.

Climate change has the potential to convert both tropical moist as well as dry forests from carbon sinks to sources.
-->

4. **Line 86: The description of vegetation sampling should be improved. It is not clear how the proportion of area occupied by each species is measured. Is it based on basal area o crown dimensions?**

We have clarified as follows:

>>>Plant cover refers to the relative projected area covered by a species [@damgaard_estimating_2014].

5. **Line 100: Are there more than one plot within the same NDVI pixel?**

There is not one pair of plots which are located in the same pixel (see Table \@ref(tab:ndvi)).

```{r ndvi, echo=FALSE, eval=TRUE, message=FALSE}
# attach packages
library("knitr")
library("dplyr")
library("raster")
library("reshape2")
# attach raster stack
ndvi = readRDS(here("images/00_ndvi_rasters.rds"))
# attach env_data
env_data = readRDS(here("images/11_env_data.rds"))
env_data$year = gsub(".*_", "", env_data$pnr)

# just keep unique plots (could have also used the id-column)
data = env_data[!duplicated(paste(env_data$lon, env_data$lat)),
                  c("lon", "lat")]
# extract values from NDVI rasters to plot locations
vals = raster::extract(ndvi[[c("ndvi_11", "ndvi_12", "ndvi_16", "ndvi_17")]], 
                       data[, c("lon", "lat")])
# count different NDVI values for each sampled year
tab = melt(vals) %>%
  group_by(Var2) %>%
  summarize(no_plots = n(), distinct_ndvi_values = n_distinct(value))
tab = rename(tab, year = Var2) %>%
  mutate(year = gsub("ndvi_", 20, year))
kable(tab, "latex", booktabs = TRUE, escape = TRUE,
      caption = "Showing that each plot has a distinct NDVI value.")
```

6. **Line 101: Worldclim data are static and describe the general pattern of variation, but I´m not sure if this database is appropriate for this study. Can you explore the inclusion on TRMM data? These data have a coarse spatial resolution but a fine temporal resolution and could improve your analysis.**

```{r gam-fig, echo=FALSE, message=FALSE, results = "asis", }
library("mgcv")
require("itsadug")

# attach response (DCA scores)
data = readRDS(here("images/15_model_input_data.rds"))
data$id = as.numeric(data$id)
data$year = as.factor(data$year)
fo = as.formula(DCA1 ~ s(ndvi, by = year, k = 3) + year)
gam_1 = gam(fo, data = data)
# gamtabs(summary(gam_1))
# plot_smooth(gam_1, view = "ndvi", plot_all = "year", ylim = c(-2, 3))
```

We appreciate the recommendation.
However, the spatial resolution of TRMM data is too low for our data, i.e., one pixel would cover several plots.
Please note also that we already have used the MODIS data for the vegetation period of each year, i.e., we have used MODIS data from 2011, 2012, 2016 and 2017. 
This also means that the NDVI for each year was computed as the mean for all NDVI MODIS images available for the vegetation period (March to May). 
We say so now more clearly in the text as follows:

>>> We downloaded the normalized difference vegetation index (NDVI) raster imagery from the Moderate Resolution Imaging Spectrometer (MODIS) available for the vegetation period (March, April) for each studied year (2011, 2012, 2016, 2017) at a 250 × 250 m^2^ resolution for our study area.
Since three NDVI scenes were available for each vegetation period, we computed the mean NDVI for each vegetation period.

Please note also that our GAM model clearly reveals a relationship between the interaction of the MODIS NDVI and year and the response variable (adjusted R squared: `r round(summary(gam_1)$r.sq, 2)`) proving that the MODIS data is appropriate for the chosen modeling approach.

7. **Line 105: This is an interesting integrative framework, but it is not clear to me what the source of species in each one plot is after the experiment was establishment. Where they planted as seeds? Is the result of immigration from nearby forest remnants or are artificially included in each plot?**

Thank you very much for the compliment!
The experiment was established on the campus of the University of Piura.
The campus comprises large undisturbed areas mirroring the typical vegetation formation, open xeric shrublands, typical of the region around Piura.
Hence, we only observed how the natural seed bank developed when exposed to artificial water/nutrient input.
Please refer also to the Supplementary Information for a detailed description of the experimental setup.

8. **Line 132: What is Londo-scale?**

Actually, we were referring to the decimal scale proposed by Londo. 
The decimal scale has finer intervals than the Braun-Blanquet scale (Table \@ref(tab:decimal-scale)).
In addition, the decimal scale does not use symbols like `r` and `+` which are hard to handle for software expecting numeric values such as ordination techniques.
We have added following sentence to the description of the vegetation sampling:

>>>Prior to the analysis, we transformed the plant cover values in accordance with the decimal scale by Londo [@londo_decimal_1976].

```{r decimal-scale, echo=FALSE}
tab = tibble(
  "coverage (in %)" = c("<1", "1-3", "3-5", "5-15", "15-25", "25-35", "35-45", "45-55", "55-65", "65-75", "75-85", "85-95", "95-100"),
  "londo_values" = c(0.1, 0.2, 0.4, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
cap = "Decimal scale as proposed by Londo (1976)."
kable(tab, "latex", booktabs = TRUE, escape = TRUE, 
      # align = "c",
      # col.names = linebreak(nms, align = "c"),
      caption = cap)
```

9. **Line 138: The modeling approach maybe inappropriate. The authors used a detrended canonical analysis to reduce dimensionality of the community matrix and then used a generalized additive model to compare the DCA axis with environmental data. There are other methods more appropriate to evaluate the effect of climate and other factors on species composition such as canonical correspondence analysis or redundancy analysis. Also, the generalized dissimilarity modeling allows to account for non-linear relationships among predictors and species composition variation.**

Apparently, we needed to rewrite our methods section to communicate more clearly how we have addressed methodologically our aims.
We have now made crystal-clear in the text that we were interested in spatially predicting the floristic gradient represented by the scores of the first DCA axis.
Kindly refer to the completely rewritten section "Modeling and predictive mapping" which now reads as follows:

>>>One aim of this study was to spatio-temporally predict ecosystem diversity (floristic composition) represented by ordination scores along the climatic gradient in the study area.
@muenchow_coupling_2013 have already shown that NDVI is an excellent predictor for vegetation assemblages in the study area.
Therefore, we modeled the scores of the first DCA axis representing the main observed floristic gradient in the study area with the help of a Generalized Additive Model (GAM) as a function of the interaction between NDVI and year:

>>>$$
\begin{aligned}
\text{Scores}_{i}\ \sim\ \mathrm{N}\left( \mu_{i},\ \sigma^{2} \right) \\
\mu_{i} = \alpha + f_{\text{year}_{i}}\left( \text{NDVI}_{i} \right) + \beta_{1}\text{year}_{i}
\end{aligned}
$$

>>>(Formula 1)
>>>
>>>where *i* refers to the *i*th observation, *$\alpha$* is the intercept, *$\beta$*~1~ is the estimated coefficient for year and *f* is a spline smoother for the NDVI:year interaction.
>>>
>>>The sole purpose of the model is the spatial prediction of the ordination scores, not statistical inference or the interpretation of the models' coefficients.
We will use variation partitioning later on to assess how the influence of environmental predictors has changed over the studied years (see also next section). 
Consequently, here we are only interested in a good predictive performance of the model to make sure that it is able to generalize, i.e., that it is able to predict reasonably well unseen data [@james_introduction_2013; @kuhn_applied_2013]. 
Only this can be the basis for a reliable map of the predicted ordination scores.
To make sure this is the case, we used spatial cross-validation.
Overall, cross-validation accounts for over-optimistic predictions on the training set by randomly splitting a data set into partitions used for training and testing [@brenning_spatial_2012;@lovelace_geocomputation_2019a]. 
Spatial cross-validation is an extension that measures a model's ability to spatially predict the response variable in the presence of spatial autocorrelation [@schratz_hyperparameter_2019]. 
We ran a 100-repeated 5-fold spatial cross-validation in which the partitioning is based on *k*-means clustering (*k* = 5; @russ_data_2010).
We also made sure that all annual observations of one plot were jointly placed in either the test or the training data set [@meyer_improving_2018]. 
We used the normalized root mean-square error (NRMSE) in percent as performance measure for the spatial cross-validation.
>>>
>>>$$
NRMSE = \frac{\sqrt{\frac{\sum_{i = 1}^{n}{(e_{i} - y_{i})²}}{n}}}{max(y) - \ min(y)}*100
>>>$$ 
>>>(Formula 2)

>>>where *n* is the number of observations, *e* corresponds to the residuals, and *y* refers to the observed values.

Since spatial prediction was the aim, we were not very much interested in the effect of the predictors (climate, NDVI) but only in a good predictive performance of the model to ensure that it is able to generalize [@james_introduction_2013].
But please note that the variation partioning uses the RDA approach, as suggested, to assess the impact of environmental predictor groups on the floristic composition across time (see also reply to your next comment).

Of course, we are also aware of canonical (constrained) ordination techniques [@legendre_numerical_2012]. 
Constrained ordination techniques model every single species as a function of environmental predictors, and then use the fitted values as input for an ordination method - either a PCA (then called RDA) or a CA (then called CCA).
One shortcoming of these techniques is that they usually explain less of the observed variance.
Another reason why we have not used constrained techniques for the predictive modeling was that we would have to use the same predictors (NDVI) twice - once as input for the CCA/RDA and then again for modeling the scores of the CCA/RDA.

<!--
One shortcoming of constrained techniques is that they exclusively use linear models, which might not be the most appropriate choice for each species encountered along a pronounced environmental gradient [@borcard_numerical_2018].
Secondly, in our case, species were recorded using the Londo scale which is a categorical variable. 
However, a linear model expects a response variable that follows a Gaussian distribution.
That alone would make the usage of a constrained ordination technique at least questionable.
Still, we could use one, and though we would not very much trust the modeled values of a canonical ordination technique for the reasons listed above, we could still use its scores as a response variable instead of the scores of a DCA.
But again, there are at least two reasons why we would refrain from such an approach:

1. Constrained techniques usually explain less of the observed variance of the original input matrix compared to their unconstrained counterparts since they only include in the ordination what the model was able to explain, the rest, i.e., the residuals, are simply ignored, which of course is not the case in the case of unconstrained techniques.
1. We would have to use the same predictors (NDVI and time) for the canonical ordination and the subsequent score modeling, i.e., we would e.g., model the CCA scores which themselves are the fitted values based on the same predictors. 
This smells very much like circular reasoning.
--> 
Finally, the generalized dissimilarity modeling approach is interesting, as it allows to model species turnover (beta diversity), and in fact we will give it a try in a future study.
However, in this study, we were interested in modeling the spatio-temporal change of the vegetation assemblages along a pronounced climatic gradient during four different ENSO years.

10. **Line 156: The variance partitioning approach used in not clear. If it is based on the GAM models employed previously, it is incorrect because they are based on the likelihood of the model (explained deviance) that is similar to R2 in an ordinary least square model, but not the same. In addition, the redundancy analysis approach (RDA) is more frequently used to assess the effect of environment on species composition. Also, the use of similar modeling methods previously employed allow comparability with other studies.**

We are sorry again that obviously we have not presented our methodological approach in sufficient detail.
First, the variation partitioning is not based on the GAM.
Instead, it is based, as suggested in your comment, on an RDA approach.
The variation partitioning computes with the help of RDA how the influence of predictor groups - in our case edaphic and topographic variables - has changed with different levels of water availability [@borcard_numerical_2018].
<!--
Second, you are right that RDA is used to assess the effect of environmental predictors on the species composition.
However, a RDA is not able to quantifiy the effects of different predictor groups - in our case edaphic and topographic variables - on species composition over time.
-->
We now made our reasoning now much clearer in the methods sections.
The updated text in the manuscript reads as follows:

>>>Variation partitioning quantifies the variation explained of the response, in our case the first two DCA axes, by one predictor group, in our case edaphic variables, while controlling for the effect of another predictor group [@borcard_numerical_2018; @peres-neto_variation_2006], in our case topographic variables.
Internally, variation partitioning computes a Redundancy Analysis (RDA) separately for each predictor group and for all predictor groups together, and subesequently quantifies how much each predictor group explains alone and jointly with the other predictor group of the variation of the response.
This is frequently displayed with the help of a Venn diagramm.
We ran the variation partitioning for each studied year (2011, 2012, 2016, 2017) to assess how edaphic and topographic variables influence vegetation composition under increasingly humid conditions.

Please note that upon acceptance we will make our analysis, data and software environment publicly available to make our research reproducible.
The variation paritioning will be available in the code repository under `code/analysis/17_varpart.R`.

11. **Line 173: The results of local species richness in Figure 2 compare among years and along the spatial gradient. But the authors do not describe in the Methods section the approach used to assess these differences? Are they using an ANOVA? Also, the modeling of local species richness maybe done in a similar way than the models employed here with DCA axes. Also, these models need to include spatial autocorrelation in their structure.**

Figure 2 is just an exploratory figure.
The line smoother was only added to aid visual interpretation.
Hence, we are not using an ANOVA here.
We are sorry for the confusion.
The caption now contains a sentence making this clear:

>>>To aid visual interpretation, loess smoothers were added.

And of course, we could have also modeled species richness, however, the aim of the study was to model the floristic gradient, i.e., the change of the floristic composition along the gradient.
Please see also our reply to comment #9 where we describe our modeling approach in more detail including how we have addressed spatial autocorrelation.

12. **Line 217: The results based on the irrigation experiment are focused on plant cover rather than plant species diversity that is the main goal of the study. These results are incongruent with the previous approach that is assessing the effect of a spatiotemporal water availability gradient on species diversity. The integrated approach is very interesting, but the response variable should be the same in the observational and experimental analyses to allow comparability. The authors can focus on plant coverage change across the plots is response to water availability. Also, if the goal is diversity, they should show results of irrigation and nutrient enrichment on the plant diversity in their experimental plots.**

In fact, there was no mismatch regarding the response variables for the predictive modeling and the irrigation-nutrient experiment. 
Sorry for not making this clear from the beginning.
We recorded the plant coverage of each species in the observational analyses and the experiments.
The decimal scale transformed plant coverage values (see also reply to comment #8) were the input for the Detrended Correspondence Analysis whose results were subsequently used as response variable for the predictive modeling approach.
In the case of the experiment, plant coverage values were simply plotted against time.
We hope we have made this clearer in the rewritten version.
Additionally, we have plotted alpha diversity against time for the experiment data (see Figure \@ref(fig:alpha-exp)).
<!-- JM: add figure to Appendix and reference it in the text --->
We have added the corresponding figure to the Supplementary Information.
Please note that alpha diversity was not very high since mostly a few dominant species (grasses) have covered the experimental plots.
This is also in correspondence with our observations from the field (see discussion section).

```{r alpha-exp, echo=FALSE, fig.cap="Number of species plotted against time during the duration of the experiment.", message=FALSE}
p_1 = readRDS(file = here("images/18_irrigation_spri.rds"))
print(p_1)
```

13. **Finally, some of these papers could be included in the introduction and discussion sections**

Thank you very much for taking the time to recommend relevant literature. 
We now cite @engelbrecht_drought_2007, @condit_species_2013, @alvarez-davila_forest_2017, @esquivel-muelbert_biogeographic_2017, @esquivel-muelbert_seasonal_2017 in our rewritten introduction and discussion sections.

# Reply to reviewer #2

1. **The manuscript addresses a very relevant question regarding our understanding of the dynamics of tropical seasonal ecosystems, especially those under strong influence of events like ENSO. The results are interesting, but I have not been able to convince myself that the approach taken by the authors is adequate. First, my main concern is that the four years of data, are not enough to reveal the relationship of ENSO with these coastal ecosystems. I agree that ENSO is a very complex phenomenon (as shown by the very heterogeneous response of precipitation in the four years of data collection, irrespective of what type of ENSO event happened) and as such it needs more intensive and long-term data. As shown by the authors, rainfall seems to be a good variable to study the response of plant communities in the area and for the time span used, but this general conclusion has been shown also by previous studies. In this regard, I would have also liked to see and discussed references addressing similar questions (e.g. the study by Elsa Cleland et al. Ecology 2013, in temperate grasslands).**

First of all, thank you very much for taking the time to improve our manuscript with your very helpful comments.
We agree that much more research is needed to deepen our understanding of the relationship between ENSO and the dry Peruvian coastal ecosystems.
Still, monitoring vegetation development across four different ENSO years along a pronounced climatic gradient is a rather unique research opportunity, which probably has not been done before.
Additionally, we would like to point out that one of our co-authors has revisited the study area each year since 1997.
Moreover, two of our Peruvian co-authors are living in Piura, one of them has even been born there.
This is to say that we are quite familiar with the dynamics of this ecosystem in response to ENSO events.
In our manuscript we have backed up this expert knowlegde by a data-driven approach.
Still, we plan to continue the revisiting of our plots within a new proposed research project, in which we will focus on how climate change affects tropical dry forests on various spatial scales.
This is of utmost importance especially when considering that climate change will affect the intensity and frequency of ENSO events.

Thank you for suggesting relevant literature.
The updated discussion pararaph now reads as follows:

>>>Beta diversity (expressed as the range of the first DCA axis), increased with wetter conditions (Fig. 2 & Table 1).
Naturally, higher species turnover was mainly driven by short-lived, less abundant species.
Concordantly, @cleland_sensitivity_2013 reports similar findings from temperate grasslands.

Please see also our reply to your next comment.

2. **Second, I would like to know what the authors think about a lagged response of the studied ecosystem to the ENSO conditions of the previous season (or seasons). Again, there are some interesting studies looking into this in temperate grasslands (e.g. Joan Dudney et al. J. Ecol. 2016, Brigitte Tenhumberg et al. Ecology 2018), and it is not improbable that similar mechanisms operate in the study area. I acknowledge that the data in the paper are not enough to incorporate this into the manuscript, but I would have welcomed its discussion.**

Thanks for pointing us to the time-lagged response of ecosystems.
Accordingly, we have added to the discussion following paragraph:

>>>Though we focussed on how current-year precipitation affects the vegetation composition, previous year rainfall might also have an effect [@tenhumberg_timelagged_2018].
For instance, @dudney_lagging_2017 observed an increase in grass and a decrease in forb abundance when rainfall was high one year earlier.
Similarly, @walter_plants_2011 suggest that grasses might have a 'drought-memory'.
Time-lagged responses can be also observed in woody species.
For example, tree ring growth of *Prosopis pallida* is highly correlated with rainfall of the previous year [@salazar_effect_2018].

3. **Third, I would have liked to see the authors use and discuss their results in the light of the ENSO index developed specifically for the southern Ecuadorean and northern Peruvian coasts to characterize conditions during each season (the Índice Costero El Niño – ICEN, see http://www.met.igp.gob.pe/publicaciones/2014/ElNino_v2_27_03_2014_7.pdf). This index seems to reflect much better the oceanic conditions in the area than the ONI index, as the latter covers a huge, nearly Pacific-wide region.**

Thanks for the recommendation.
In fact, we would love to use a local index.
However, the proposed index, does not classify any year between 1999 to 2017 as La Niña (when considering the three-month mean of December/January/February and January/Feburary/March) which strikes me as a bit odd (see Figure \@ref(fig:bp)).
For example, 2011 and 2012 have been La Niña years. 
And La Niña had a strong influence on the coastal vegetation of Peru as we clearly observed on Lomas mountains near Casma and Trujillo [@muenchow_predictive_2013a;@muenchow_soil_2013].
It is also interesting that 2005 was not classified as La Niña despite much lower rainfall than average across the complete study area.

```{r bp, echo=FALSE, message=FALSE, fig.width=12, fig.height=4, fig.cap="Ratio between annual precipitation and median precipitation (median Paita: 25.9mm, median Piura: 71.4mm, median Chulucanas: 272.4mm) measured at the three automated climatic stations (see section 2.2.2). In case the precipitation was smaller than the median, we computed the ratio as: 1 / (precipitation / median(precipitation)) * -1. The years in which vegetation sampling took place are printed in bold (x-axis). The ENSO phases are in correspondence with the ICEN index found at: http://www.met.igp.gob.pe/datos/icen.txt."}
# the result is the barplot figure (bp)
source(here("code/misc/explore_icen_index.R"))
print(bp)
```

The index seems to be interesting but still is a bit experimental as also pointed out in the publication you have indicated.
Therefore, we have refrained from using the ICEN index but we have prominently mentioned it in a rewritten discussion section now named *ENSO, climate change and ENSO indices*:

>>>Naturally, local anomalies possibly overlap, dilute and modify the ENSO signal stemming from the central Pacific. 
Studies have shown that indices computed from the El Niño region 1+2, which corresponds to an area in the tropical Pacific close to the South American coast, already provide better correlations with climate anomalies in northern Peru and southern Ecuador [@kao_contrasting_2009;@salazar_effect_2018]. 
At least two further local indexes seem to be also promising for predicting climate anomalies along the Peruvian coast.
The first is the Índice Costero El Niño (ICEN; see http://www.met.igp.gob.pe/publicaciones/2014/ElNino_v2_27_03_2014_7.pdf), and the second is an ENSO index for the insurance of agricultural products [@mortensen_enso_2018].

4. **Fourth, I do not agree with several of the statements made in the Conclusions. To begin with, I do not think that there is a current simplistic understanding of ENSO. For several years now, government and academic sectors working in southern Ecuador and northern Peru are officially using locally developed indexes with a complex array of ENSO definitions and characterizations. And I expect this to be reflected with increased presence in the scientific literature in coming years.**

We absolutely agree.
Hence, we have completely rewritten our Conclusions:

>>Constant monitoring is a prerequisite for a better understanding of ENSO-affected ecosystems. 
Since ENSO exerts its most direct influence on tropical regions, and all corresponding ecosystems are endangered without exception, the correct prediction of the expected ENSO effects are a cornerstone of informed conservation and reforestation management. 
Conserving and reversing the state of degraded ecosystems also helps to mitigate the effects of floodings and landslides due to excessive rainfalls, and thereby contributes to protect the population and infrastructure from unnecessary harm [@fraser_peru_2017; @ramirez_understanding_2017].
Currently, the regional government has set up eleven restoration projects in northern Peru which are trying to protect the endemic vegetation and increase vegetation cover through reforestation [@cordova_piura_2019].
Additionally, NGOs and mining companies are also working on reforestation projects [@gutierrez_anuncian_2015].
Most of these projects rely on extreme events (like the 2017 El Niño) to boost plant growth to overcome the otherwise very limited access to water.
Our experimental results indicated that water availability showed a non-linear relationship with biodiversity and plant cover.
Therefore, restoration projects might consider water regulation in combination with a sensible amount of fertilization [@dahlsjo_local_2019].
The same applies to a sustainable agrarian management.
Indeed, agricultural development has been very intense over the last 20 years in the study area.
Personal communications with the main local companies (e.g., https://www.ecosac.com.pe/) indicated that extreme events are an important issue.
On the one hand, El Niño has the potential to boost crop productivity.
On the other hand, especially short and very intense rainfall might create havoc during the crop season as was the case during the 2017 El Niño.
Additionally, our results clearly reveal patterns in the spatial variability of biomass production in response to water availability along the climatic gradient, which might be also an important information for agrarian and restoration activities.

5. **Furthermore, after reading this section several times, it is not clear to me how the present study can achieve the statements presented in the Abstract and barely touched upon in the rest of the manuscript and in this final section in particular. The abstract mentions that the results from this study can help “protect local human populations from devastating ENSO effects through informed conservation management and restoration activities, and that the study can help equip farmers with information on sustainable agrarian management”. I agree that both are valid and important points to consider, but I would have preferred the authors to expand on this and provide logical connections between their results and restoration ecology approaches in seasonal tropical environment and TDFs. Likewise, I would think that the high-tech agricultural development that has occurred in the study area in the past 20 years has produced knowledge in terms of dealing with variable climatic conditions and I would have expected some literature review on this as well.**

Again, we absolutely agree and we appreciate the reviewer's concerns about the expected impact of our results.
Therefore, we have completely rewritten the conclusion section which now, as suggested, focusses much more on restoration ecology approaches (see also reply to your previous comment).
Unfortunately, we could not find any peer-reviewed scientific literature on high-tech agriculture from North Peru, and therefore refer to personal communications we had with local business and governmental stakeholders.

6. **29: why is the area one of the most climatically vulnerable regions on earth?**

Thanks for noting, we toned down as follows while also offering a reference:

>>> one of the most variable regions on Earth in terms of climate [@wehrden_global_2012].

7. **30: I have never seen the term “endemic” applied to an ecosystem. How do you define an endemic ecosystem?**

We replaced *endemic ecosystem* by *ecosystem featuring a high rate of endemism*.

8. **31: I beg to disagree, but TDF almost as high in biodiversity as humid tropical forests? In what terms?**

We agree this is indeed surprising but it is what the literature says:

> "Tropical dry forests have smaller stature and lower basal area than tropical rain forests (Murphy & Lugo 1986), they are diverse both locally and regionally (Givnish 1999) and although less known, they are almost as diverse as tropical rain forests."

@espinosa_what_2011

> "There is also a general assumption that rain forest is the biome richest
in species in the tropics. However, a recent evaluation of plant species diversity in dry forests across Latin America and the Caribbean suggests that they have more than 7500 woody species, almost certainly an underestimate because it is based on only 1600 inventories covering 125,000 square km across this vast region (for example, the dry forests of NE Brazil cover 750,000km^2 alone; Figure 1). This suggests that in their entirety, the dry forests of the New World may
contain as many or possibly more species than the Amazon rain forest, where the most recent estimate of tree species numbers is ~6200. Similarly, considering all plant growth forms, the savannas of Central Brazil contain 11.384 species, 35 more than the Brazilian Amazon rain forest."

@pennington_tropical_2018

9. **34: TDFs are understudied, and a specific reference for this statement would also be Escribano-Avila et al. 2017**

Thanks! 
We have added @escribano-avila_biodiversity_2017 to the text.

10. **37-39: the group of Carlos Espinoza from UTPL in Ecuador has conducted several studies looking into these questions. There are also quite a few studies in Mexico and Colombia, specifically in TDFs. I would consider some of those, key references for the manuscript.**

Thank you very much for the literature recommendation.
We now additionally cite @ulrich_climate_2014 in the introduction.
Additionally, we have added following sentences to the discussion section:

>>>However, the most important topographic variable is altitude since it is highly correlated with mean annual precipitation in the study area. 
This is also in correspondence with results from Mexico where topography played also a part in explaining the composition of dry forests [@mendez-toribio_effects_2016].

and ...

>>>Other studies have also confirmed the importance of soil for the composition of the vegetation in drylands.
For example, @ulrich_climate_2014 found that soil fertility and variability in rainfall and temperature are the most important predictors for beta diversity.
Similarly, soil plays an important role in Colombian dry forests [@gonzalez-m_disentangling_2018] 
Finally, @espinosa_what_2011 found out that soil moisture, soil temperature and nitrogen explained in large parts the floristic composition of Tumbesian dry forests, which is quite close to our study area.

11. **80: “uniform during normal years”. Based on the previous text, what is considered normal? No rains or even drought?**

We are sorry, of course, we were referring to neutral years, i.e., neither La Niña nor El Niño years.
We have updated the text accordingly.

12. **82: Were plants collected? How was the ID process? In my experience, even after years of monitoring permanent plots in these seasonal environments, there is a percentage of unidentifiable material, especially if it is collected during the vegetative growth phase.how did you ensure across-season uniformity in species ID**

We collected all plants which were unidentified in the field.
Our co-author, Eric Rodríguez, who is a botanist and leading the herbarium at the University of Trujillo, determined the unidentified species.
Still, a few species remained for which either the genus nor the species could not be determined.
These were only named after their family, e.g., unkn_Sola for unknown Solanaceae.
If more than one species from the same family could not be identified, we simply numbered the names, e.g., unkn_Sola1 and unkn_Sola2.
If even the family could not be determined, the plant was identified as unknown.
To be able to recognize the same species, especially the unidentified species, during each sample  we have taken various fotos of each single species.
The data sampled in the field was stored in an Access database using specifically designed forms to ensure that the same attributes for each plot were recorded and that the same species names were used for each plot.

```{r unknown-table, echo=FALSE}
pa = readRDS(here("images/11_ct_pa.rds"))
d = data.frame(unidentified_species = sort(grep("unkn", names(pa), 
                                                value = TRUE)))
cap = paste("Unidentified species. Please note that unkn\\_Faba1 was", 
            "determined later on which is why it disappeared from the table.")
kable(d, "latex", booktabs = TRUE, escape = TRUE, 
      # align = "c",
      # col.names = linebreak(nms, align = "c"),
      caption = cap)
```

13. **83: What do you mean by “randomly sampled”? You randomly sampled from 50 plots? Or do you randomly established 50 plots along the gradient?**

We have clarified as follows:

>>> We randomly sampled 50 vegetation plots of 30 × 30 m^2^ along the climatic gradient between Paita and Chulucanas (Fig. 1).
To ensure that all parts of the study area were equally likely to be sampled, we stratified the sampling by distance to the Pacific Ocean in ten sampling zones.

14. **84: what are those 10 classes? They are not mentioned anywhere in the text or figures.**

As mentioned in the previous reply, the ten classes are just zones stratified by distance to the sea as shown in Figure \@ref(fig:sampling-zones).
We have added the figure to the Supplementary Information and referenced it in the main document.

```{r sampling-zones, echo=FALSE, fig.cap="Study area stratified by distance to sea in 10 classes."}
p_1 = readRDS(here("images/18_sample_stratification.rds"))
print(p_1)
```

15. **84: that whole study area is a mosaic of different land uses and scarce (if any) areas devoid of human use. Humans and especially free-ranging cattle could turn into major disrupting factors of vegetation dynamics. Was this considered when selecting location of plots, and more importantly, when evaluating results?**

You are absolutely right.
The humidity gradient is at the same time a land use gradient.
Please refer also to the methods section in which we describe in detail how plot locations were sampled.
Prior to the sampling, we have excluded explicitly urban, riparian and agricultural areas. 
The latter included mostly farmed fields but also fenced areas to enclose goats, and rarely also cows and horses.
Still, some of these animals are allowed to roam freely the properties of their owners.
Therefore, we added a sentence to the methods...

>>>Agricultural activity includes field farming and extensive grazing caused by free-ranging livestock.

... and to the discussion section:

>>>Finally, free-ranging livestock might also have an impact on vegetation dynamics in the study area, especially in its eastern parts with more than 100mm of annual precipitation.
Livestock might alter floristic composition and diversity while supporting especially plants adapted to grazing.

16. **90: Did you sample soils only once or for every field season? If once, any idea if soil composition or chemistry could vary between years?**

We collected soil samples only once, and therefore have updated the text as follows:

>>>We randomly took and mixed three soil samples from 15cm and 30cm depth in each plot in 2011. 
Soil composition should not vary too much between the years in which vegetation sampling took place since the study area largely features arenosols, i.e., a very weekly developed soil type.
On the other hand, soil chemistry can vary largely between two years.
For instance, El Niño events can increase soil respiration, i.e., soil CO\textsubscript{2} emissions, by a factor of 100, especially in the vicinity of trees [@salazar_tree_2019].

17. **99: the vegetative period is from-to?**

The vegetation period usually lasts from March to May, we now also say so in the text.

18. **132: what is the Londo-scale?**

Please refer to our reply to comment #8 of reviewer #1 who also raised the same question.
We are sorry that we have not made clear from the beginning that we are referring to the decimal scale created by Londo [@londo_decimal_1976].

19. **174: results from Table 1, I would have liked to see some max-min ranges of each variable.**

Thanks for the advice. 
We have added the requested columns to the table.

20. **256-284: Very interesting discussion, and I wonder what differences might be revealed if the data is stratified by growth form (annual herbs vs perennial woody species for instance).**

```{r lf-dca, eval=TRUE, echo=FALSE, fig.cap="DCA species plot with the species colored by lifeform.", out.width="75%", fig.align="center", message=FALSE}
# attach packages
library("vegan")
library("dplyr")
library("lattice")
library("magrittr")
library("ggplot2")
library("RSQLite")
# source own functions
source(here("code/funs/ordi_helper_funs.R"))
source(here("code/funs/bestisomap.R"))
# attach data
londo = readRDS(here("images/11_ct_londo.rds"))
cover = readRDS(here("images/11_ct_cover.rds"))
env_data = readRDS(here("images/11_env_data.rds"))
lf = readRDS(here("images/11_condensed_lifeforms.rds"))

# dca = ord_fun(londo, env_data, method = "dca")
dca = decorana(dplyr::select(londo, -pnr), iweigh = 1)

sc = scores(dca, display = "species")
# just checking
# setdiff(rownames(sc), lf$species)  # 0, perfect
# setdiff(lf$species, rownames(sc))  # 0, perfect
sc = as.data.frame(sc)
sc$species = rownames(sc)
sc = inner_join(sc, lf, by = "species")
cols = c("darkgreen", "yellow3", "brown")
p_1 = xyplot(DCA2 ~ DCA1, data = sc, 
             groups = sc$type, col = cols, pch = 19, 
             auto.key = list(corner = c(0, 0)),
             par.settings = 
               list(superpose.symbol = list(col = cols, pch = 19)),
             scales = list(alternating = c(1, 0),
                           tck = c(1, 0)))
print(p_1)
# # another idea would be to use the percentage of each lifeform for each plot
# # as predictors in envfit
# # run through plots, merge lifeform and compute percentage of each lifeform
# lf_agg = plyr::ldply(seq_len(nrow(cover)), function(i) {
#   tmp = cover[i, ] %>%
#     reshape2::melt(.)
#   id = tmp$pnr[1]
#   tmp = inner_join(tmp, lf, by = c("variable" = "species"))
#   tmp_2 = group_by(tmp, type) %>% summarize(portion = sum(value)) %>%
#     mutate(percentage = portion / sum(portion) * 100) %>%
#     select(-portion) 
#   tmp_2 = melt(tmp_2) %>% dcast(., variable ~ type)
#   tmp_2 = rename(tmp_2, id = variable)
#   tmp_2$id = id
#   # return your result
#   tmp_2
#   })
# fit = envfit(dca, select(lf_agg, -id))
# 
# # retrieve arrow heads manually, see here:
# # browseURL("https://stackoverflow.com/questions/14711470/plotting-envfit-vectors-vegan-package-in-ggplot2")
# 
# coords = scores(fit, "vectors") * vegan:::ordiArrowMul(fit)
# plot(dca, display = "sites")
# plot(fit)  # coords correspond to arrowheads 
# # even better would be to compute the envift for each year but then you would
# # have to run a DCA for each year individually

```

Thank you for the compliment!
And yes, you are right, it would be interesting to see if there are differences along the gradient between lifeforms.
Fig. \@ref(fig:lf-dca) is a first attempt to delve deeper into this question.
However, for the moment we chose not to discuss the implications further in the manuscript since the editor asked us to stick to the most important results.

21. **368: Mention of Panel b is actually c?**

Thanks for spotting.
We changed the caption accordingly.

# Bibliography