### 2020-08-10 text/ecography/round_3/reply_to_reviewers_2.Rmd

---
output:
  bookdown::pdf_document2:
    keep_tex: no
    latex_engine: xelatex
    # float is needed for floating figures
    extra_dependencies: ["fontspec", "sectsty", "float"]
    toc: false
    fig_caption: yes
    number_sections: true
header-includes:
  # adjust figure captions
  - \usepackage{caption}
  - \captionsetup{font={footnotesize,bf,singlespacing}}

title: "Point-by-point reply to reviewers' comments"
bibliography: ../transect_refs.bib
csl: ../ecography.csl
# biblio-style: "natbib"
link-citations: true
---

```{r setup, include=FALSE}
library("knitr")
library("kableExtra")
library("tidyverse")
library("here")
library("RSQLite")
library("sf")
# make sure that figures are allowed to fload (hold the assigned position and are not always put on top of a page)
knitr::opts_chunk$set(fig.pos = 'H')
# needed for dates in experiment figure species richness, otw months are displayed in German
Sys.setlocale("LC_TIME", "C")
```

# Reply to subject editor Bethany Bradley

Dear Dr Bradley,

thank you very much for handling our manuscript.
We are pleased to resubmit a revised version of our manuscript entitled  *Monitoring and predictive mapping of floristic biodiversity along a climatic gradient in ENSO's terrestrial core region, NW Peru*.
We are also very thankful for the constructive comments of both reviewers which helped to improve substantially our manuscript.
Please find below a point-by-point reply to all concerns raised by the reviewers.
In accordance with your and the reviewers' suggestions, we have:

- integrated into a coherent story the vegetation transect analysis and the irrigation-fertiization experiment.
- deleted parts of the discussion to discuss above all findings supported by our data.
- completely rewritten the Abstract, the Conclusions and in great parts the Introduction to make our aims and the corresponding results clearer.
The Conclusions now exclusively feature the 'take-home' points providing an answer to the hypotheses posed in the introduction. 
- added a new figure showing how species richness developed in the experimental study.

After having addressed thoroughly and carefully the reviewers' comments for the third time, we now hope that you deem our manuscript suitable for publication in *Ecography*.
In any case, we would like to express our gratitude again towards you and the reviewers for taking the time to consider and revise our manuscript.

Yours sincerely,

The authors

# Reply to reviewer #1

1. **The different studies are not well combined and appear too much in parallel, especially in the discussion section. Also it is unclear why the spatial prediction was performed because it does not appear to be needed to discuss the findings. In general, the manuscript needs a lot more clarification in the presentation of the aim, hypotheses, and discussion of the results.**

Thanks for taking the time to revise and improve our manuscript.
We have integrated into a single story our transect analysis and the water-fertilizer experiment.
Please refer to the replies to comment #7, 11, 23.

The spatial prediction is crucial for this study.
We have added a new paragraph to the introduction, described the reasoning in more detail and highlighted the resulting findings more prominently throughout the whole manuscript. 
Please refer also the replies to comments #16-19.

Furthermore, we have rewritten our hypotheses, clarified the general aim of our study, have deleted parts of the discussion and have completely rewritten our Conclusions mirroring now the posed hypotheses, and hence the structure of the manuscript.
Please refer to the replies to comments #7-14, 23, 26.

2. **The abstract is not very clearly written and is not presenting a good summary of the paper. E.g. the first two sentences are a poor motivation for this study (the motivation is to study TDF only because they are understudied?). It would be better to start with a motivation that highlight the relevance, e.g. that these areas feature unique biodiversity but that the impact of ENSO on biodiversity patterns is still unknown.**

Thanks for the advice.
We have rewritten as follows:

>Tropical dry forests (TDF) are heavily shaped by the El Niño Southern Oscillation (ENSO) in NW Peru, where especially El Niño episodes bring rain to arid to semi-arid areas.
However, the resulting effects on biodiversity patterns remain largely unknown as well as the effect of environmental variables on the floristic composition under varying rainfall patterns.
Therefore, we studied the spatio-temporal effects of different ENSO episodes on floristic biodiversity along a climatic gradient ranging from the coastal desert to the Andean foothills.
We sampled 50 vegetation plots in four years representing different ENSO episodes.
To highlight the spatio-temporal changes in floristic composition and beta diversity across ENSO episodes, we predicted ordination scores with a Generalized Additive Model.
We applied variation partitioning to test if topographic or edaphic variables gained in importance during more humid ENSO episodes.
Additionally, we executed an irrigation-fertilization experiment to quantify the beneficial effects of the water-nutrient interaction under different simulated ENSO rainfall scenarios. 
Plant species richness increased under humid conditions during the humid La Niña (2012) and the moderate El Niño episodes (2016), and slightly decreased under the very humid conditions during the coastal El Niño episode (2017).
The spatial prediction revealed that specific vegetation formations became more pronounced with increasing water input, but that a large water surplus led to the disruption of the strict order along the climatic gradient.
Edaphic and topographic variables gained in importance with increased water availability (2012 & 2016), however, this effect was not further amplified under very wet conditions (2017).
The experiment showed that plant cover under Super Niño conditions was three times higher when fertilized.
Overall, our spatial predictions concede detailed insights into spatio-temporal ecosystem dynamics of TDF in response to varying rainfall caused by different ENSO episodes while the results of the experiment can support farmers regarding a sustainable agrarian management.

3. **Carefully check for logical reasoning in the abstract. E.g. you write that you wanted to test the effect of other variables when water is no longer the limiting factor. But it was not mentioned before that water is a limiting factor.**

Thanks, please refer to the reply to the previous comment.

4. **Your aims and hypotheses other than just studying biodiversity in the TDF (which is quite a broad aim) is not getting clear. Be more specific here on what you want to find out with your study and what you expect.**

Please refer to the reply to comment #2.

5. **It appears unclear at that point how the results of this study “can help the local population from devastating ENSO effects”.**

We have toned down by rewriting the text as follows:

>Overall, our spatial predictions concede detailed insights into spatio-temporal ecosystem dynamics of TDF in response to varying rainfall caused by different ENSO episodes while the results of the experiment can support farmers regarding a sustainable agrarian management.

6. **Line 11-15: As your study area is in Peru you should mention what effect the ENSO phases have for your study area instead of a global description of the phenomenon.**

We have rewritten as follows:

>ENSO is a recurrent climate phenomenon causing anomalies around the globe [@capotondi_understanding_2015].
ENSO consists of alternating episodes.
El Niño represents the warm and La Niña the cold episode of this cycle. 
ENSO neutral refers to periods when neither El Niño nor La Niña episodes are present. 
The study area is part of the core region where the ENSO phenomenon exerts its greatest terrestrial influence [@richter_monitoring_2005]. 
Generally, El Niño episodes bring more rain than usual and La Niña episodes enforces the already dry conditions in the study area [<50 mm of annual precipitation along the coast of NW Peru; Fig. 1D;  rollenbeck_climatic_2015].

7. **In general the aim as well as hypotheses are not getting very clear in the introduction. E.g. Line 40: “To tackle the outlined issues...” Be more specific here what issues you’re talking about. In the sentences above you talk about the effect of different soil properties. But you are not addressing this in your methods. Also it’s not getting clear why you want to study edaphic and topographic properties on floristic composition. How does this link to your major aim to study bidoversity with different ENSO phases? This needs a lot more clarification. Also the motivation why different ENSO phases are studied is not getting clear. Make clear why this is relevant.**

Thank you for pushing us to:

- make our aims and hypotheses clearer.
- clarify why it is important to study biodiversity during different ENSO episodes.

We have added several new sentences to the introduction.

With regard to different ENSO episodes, we added:

>Therefore, one objective of this manuscript is to study the effect of different ENSO episodes, including La Niña episodes, on the floristic composition, comprised of all observed vascular plants, of the same ecosystem.

With regard to the predictive mapping, we added a new paragraph (see also the reply to comment #15) while already mentioning the corresponding aim as follows:

>We aim at spatially predicting, for the first time, how different ENSO episodes might change and reshape the spatial distribution of the floristic composition in the study area.

With regard to the impact of environmental variables, we added:

>Hence, this study investigates how much the impact of edaphic and topographic variables varies in explaining the floristic composition between dry and humid ENSO years.
Additionally, an irrigation-fertilization experiment should help to quantify in more detail the beneficial effects of the water-nutrient interaction on vegetation diversity and cover.

Additionally, we have rewritten the aims of our manuscript as follows:

>In summary, in this manuscript we study the effects of different ENSO episodes on biodiversity patterns and floristic composition  as well as the effect of environmental variables on the floristic composition under varying rainfall patterns.
Specifically, we tested the following three hypotheses:
>
(1) We investigated the change in the different levels of biodiversity along a pronounced climatic gradient across different ENSO episodes. 
Overall, we expected a boost in biodiversity as soon as water was no longer the predominant limiting factor.
We were particularly interested in how different ENSO episodes modify the spatial distribution of the floristic composition.
Overall, we expected a more pronounced development of vegetation formations along the climatic gradient with wetter conditions [@engelbrecht_drought_2007].
>
(2) Water availability limits plant cover and species richness along the entire climatic gradient in the (semi-)arid study area.
Therefore, we were interested in evaluating the effects of soil characteristics and topography changed in shaping floristic composition and vegetation cover (biomass production) with varying rainfall patterns as caused by different ENSO episodes. 
We expected that the influence of topographic and edaphic variables might gain in influence in wetter years, i.e., as soon as water is no longer the limiting factor.
>
(3) To quantify the beneficial effects of the water-nutrient interaction on vegetation cover development and species richness, we executed an irrigation-fertilization experiment simulating different ENSO rainfall scenarios.
We expected that the water input of the rainfall scenario representing a Super-Niño episode in combination with the fertilizer would be most conducive to plant cover and species richness development.

As now written in more detail in our hypotheses, we also wanted to test if the impact of edaphic and topographic variables change as soon as water is no longer the limiting factor in the study area.
Different ENSO episodes create different rainfall scenarios, and therefore provide an excellent opportunity to test if and how the influence of environmental variables change the floristic composition and biodiversity.

Please also note that we describe in section *2.2.2 Environmental variables* which edaphic variables were sampled in the field.

8. **Line 45/46: Don’t mix up hypothesis and methods here**

Please refer to the reply to the previous comment and to comment #18.

9. **Line 50/52: Why do you expect that topography/soil properties have a higher effect on biodiversity in wetter years? The hypothesis is not getting clear.**

Please refer to the reply to comment #7.

10. **Line 53-57: This is a description of the method and unclear what the corresponding aim is. Better describe the hypothesis/aim here and describe the methods applied to approach this later on.**

Please refer to the reply to comment #7.

11. **In the introduction it is unclear what kind of biodiversity you’re studying (not even mentioned that you focus on vegetation!). But that’s relevant for the understanding/hypotheses. E.g. for trees, I would not expect that biodiversity changes during ENSO cycles. So please state what you’re studying and why you expect changes.**

You are absolutely right.
We have added following sentence to the introduction to make clear what we study:

> Therefore, one aim of this manuscript is to study the effect of different ENSO, including La Niña, episodes on the floristic composition, comprised of all observed vascular plants, of the same ecosystem.

Please refer also to the reply to comment #7.

12. **Section 2.2.1: You often talk about plant development. Be more specific here. Do you mean biomass/coverage/diversity?**

We have clarified as follows:

>In addition to the plot survey at the landscape scale, we conducted a small-scale irrigation-fertilization experiment to quantify the beneficial effects of the water-nutrient interaction on species richness, plant cover and biomass production.

13. **2.2.1: Experimental in addition is great to test effect of water availability in a more systematic way than under real-world conditions. But was it vegetation on the experimental plots from the same ecosystem? Or how can the two surveys be compared?**

We have clarified as follows:

>We established the experiment in the middle of the study area on on a protected dryland forest inside the campus of the University of Piura (5°10'S, 80°38'W, see Piura in Fig. 1).
Hence, the experiment was located in the same ecosystem as the landscape study (see previous paragraph) and more specifically in the middle of the observed precipitation gradient, ranging from 26 mm (Paita; annual median precipitation) to 272 mm (Chulucanas; Fig. 1).

Please refer also to Appendix 2 for a more detailed description of the experiment.

14. **Section 2.2.2: Because the hypotheses/aims are not well described in the introduction, it is not clear now why soil samples were taken.**

Please refer to the reply to comment #7.

15. **Section 2.2.2: Why did you use NDVI from MODIS (250m resolution) when you have much smaller plots? Wouldn’t something like Landsat/Sentinel-2 be more appropriate? Also see comment above: Not clear at this point why these variables were calculated because the aims/hypotheses were not made clear. Also at the beginning of the method chapter a short summary would help gaining understanding of what you did for which purposes before going into detail here.**

We have not used Landsat/Sentinel-2 imagery due to following reasons:

- Sentinel-2 data have only become available globally since March 2017.
For consistency reasons, we believe it is better to use the same sensor for each modeled year instead of mixing data from different sensors for different years.
- MODIS data has a much better temporal resolution than Landsat imagery, i.e., as opposed to MODIS data, Landsat data was not available for the vegetation period in each sampled year whereas there were thress MODIS images available for each vegetation period.
Please note also that there is not one pair of locations that share the same MODIS pixel.

16. **2.3.2: Here you mention for the first time, that one aim is to do spatio-temporal prediction. This should have been already mentioned in the introduction. Also there please make clear WHY this is an aim of your study. Predictive mapping even occurs in the title but it is not made clear why you’re doing this, other than  for visualization as you mention in the introduction.**

You are absolutely right, we should have already emphasized in the introduction the importance of predictive mapping, and why we are doing it.
Therefore, we have added following paragraph to the introduction:

>Statistical learning has a long tradition in biodiversity research for detecting patterns in ecological data and predicting from these patterns [@guisan_predictive_2000].
It is especially useful for quantifying species-environment relationships and for predicting the occurrence of species or species richness (predictive mapping), and therefore is an indispensable instrument for biodiversity management [@fremout_mapping_2020].
Furthermore, predictive mapping can reveal ecological changes in space and time which otherwise might be hard or impossible to detect [@lovelace_geocomputation_2019a].
Therefore, it is a crucial tool for gaining detailed insights into ecosystem dynamics, which other more descriptive methods such as global biodiversity measures and ordination techniques cannot capture.
We aim at spatially predicting, for the first time, how different ENSO episodes might change and reshape the spatial distribution of the floristic composition in the study area.

17. **2.3.2: By using NDVI as the only proxy for assemblages, all you do is to assume that diversity increases with vegetation cover and/or biomass (as the NDVI is increasing with cover/biomass but not with diversity if the biomass is the same). Probably there is a good correlation between cover and diversity but this scientific explanation should be made clear here. However, in line 256-257 you write that 2007 was wetter and biomass increased (consequently increased NDVI) but diversity decreased. How does this fit to your NDVI~diversity model? **

In fact, we have modeled the first axis of the DCA as a function of the interaction NDVI:year.
The input for the DCA was a plot-species matrix consisting of the decimal-scale transformed cover values, hence the resulting scores comprise both cover and diversity (number of species), and represent the floristic composition of the observations condensed to a single number.
Therefore, a score can also represent a plot whose number of species has decreased while the overall plant cover has increased.
In fact, Fig. 2 in the manuscript provides a visual representation how the position of one plot can change in ordination space with changing cover and species richness values.
Please refer also to the reply to the next comment in which we included a figure that shows the relationship between the response, i.e., the first DCA axis, and NDVI by year (Fig. \@ref(fig:dca-ndvi)).
Thank you again for advising us to include this figure which is crucial since it makes transparent the observed relationship between the scores of the first axis and NDVI and how this relationship changes between years.

18. **Also, why do you include the year as predictor? By doing so, you don’t set up a model that would be in any way applicable beyond the years you sampled. Why are you not assuming that the relationship between NDVI and diversity stays the same independently of the year? **

We included the year-NDVI interaction because we wanted to allow the model to have a different intercept and a different slope for each year (Fig. \@ref(fig:dca-ndvi)).
Additionally, we used a spline smoother to allow for the observed non-linearity (Fig. \@ref(fig:dca-ndvi)).
And you are absolutely right that it is crucial to give the reader the opportunity to have a look at the relationship between NDVI and the response variable.
Therefore, we have also put Fig. \@ref(fig:dca-ndvi) into the Supplementary Information and have cross-referenced it in the main text.

```{r dca-ndvi, echo=FALSE, message=FALSE, fig.align="center", fig.cap="Relationship between the response (DCA1) and NDVI by year. To aid visual interpretation, line smoothers were added.", out.width="70%"}
data = readRDS(here("images/15_model_input_data.rds"))
ggplot(data, aes(x = ndvi, y = DCA1, group = year)) +
  geom_point(aes(color = as.factor(year)), alpha = 0.4) + 
  labs(color = "year") +
  xlab(label = "NDVI") +
  ylab(label = "DCA1") +
  geom_smooth(aes(color = as.factor(year)), se = FALSE) + 
  theme_bw() +
  coord_fixed(ratio = 0.15)
```

The aim of our study was to visualize by means of predictive mapping how the floristic composition changed between years, i.e., how different ENSO phases shape the spatial distribution of vegetation assemblages.
Therefore, we were not interested in predicting unobserved or future years.
But if we were, it would be possible to do so by taking a mean intercept and a mean slope.
However, doing so, would cancel out the majority of the interannual differences.

To make our aims clearer, we have rephrased the corresponding hypothesis as follows:

(1) We investigated the change in the different levels of biodiversity along a pronounced climatic gradient across different ENSO episodes. 
Overall, we expected a boost in biodiversity as soon as water was no longer the predominant limiting factor.
We were particularly interested in how different ENSO episodes modify the spatial distribution of the floristic composition.
Overall, we expected a more pronounced development of vegetation formations along the climatic gradient with wetter conditions [@engelbrecht_drought_2007].

19. **More general question: why do you need the spatial prediction? Isn’t it sufficient to look at the plots that were sampled to test your hypotheses?**

Only the predictive mapping approach reveals how different ENSO episodes changed the spatial distribution of floristic composition and diversity.
An ordination by itself only reveals that diversity has changed along the gradient but cannot reveal spatial patterns.
But thank you for pointing us to formulate our hypotheses and aims in more detail.
Please refer also to the replies to comments #16, 18 and to the reply to comment #4 of reviewer #2.
And please note also that we discuss the outcome of the spatial prediction in detail in section *4.2 Floristic gradient mapping across ENSO episodes*.

20. **Numbers in Fig.2 not readable**

We now show only a few plot numbers which should illustrate how the position in ordination space has changed for a few selected examples throughout the years (Fig. \@ref(fig:dca)).

```{r dca, echo=FALSE, fig.cap="Plotting the scores of the second DCA axis against the scores of the first DCA axis. To better visualize the cross-annual changes, the scores of all years were added in light gray to the background of each panel. The black dots refer to the scores of the year given in the strip of each panel. Numbers above the dots are the plot IDs which were ordered in accordance with the distance to the sea, i.e. plot number 1 is closest and plot number 50 farthest away from the sea. To avoid cluttering the figure with too many numbers, we have only added a few selected plot numbers.", out.width="100%", fig.align="center"}

library("lattice")
load(file = here("figures/14_londo_dca.rda"))
print(p_1)
```

21. **Figure 4: Scatter plot NDVI~richness would be helpful in addition as this represents the basis for the spatial prediction. Also it is unclear to me what the numbers (and colors) mean. What are the negative and positive values? According to the text below it appears to me as if the number correspond to vegetation types. But if so, I don’t understand why its expressed as a gradual color scale and with positive and negative numbers.**

The scores of a DCA axis are of type numeric, and hence can take on any value between + and - infinity.
Since the score of a DCA can be understood as the floristic composition condensed to a single number, we have interpreted certain DCA value ranges as belonging to a specific vegetation formation.

We added following explanation to the corresponding results section:

>Since the predicted DCA scores represent the floristic composition of a pixel, clusters of pixels sharing similar values (visible as similar color tones in Fig. 4) can be understood as belonging to the same vegetation formation.

22. **Figure 5: I think the indication of the year is missing (at least I assume the 4 plots correspond to the 4 years)**

Thanks for spotting.
The year labels should have been there in the first place.
Changed as requested.

23. **Also in the discussion/conclusion it remains unclear what the study is exactly about. Change in vegetation patterns due to ENSO, improved agriculture, nature conservation? The claimed implications are often not justified by data. e.g. you have higher diversity when more rain is there but does this have implications for nature conservation as claimed in the paper?**

We have rewritten the Conclusions in such a way that they now mirror the hypotheses and therefore also the overall structure of our paper, i.e, they now present the most important result for each hypothesis and draw a general conclusion from the chosen integrated approach.
We also made clear how our results can be of use in nature conservation and improved agriculture at the same time.
Please note also that we have deleted a complete section in the discussion to also streamline the discussion with the posed hypotheses (see also the reply to comment #26).
The Conclusions now read as follows:

>Alpha and gamma diversity increased with wetter conditions during the humid La Niña (2012) and moderate El Niño episodes (2016), though, a few dominant species largely prevailed under the very humid conditions of the coastal coastal El Niño episode (2017), which then even resulted in a slightly decreased alpha diversity.
Still, plant cover was greatest by far during the most humid conditions.
Beta diversity also increased with wetter conditions but did not change significantly between years in accordance with the test for multivariate homogeneity of groups dispersions.
As useful as such global measures are in general, they are in essence a single descriptive figure which naturally lack the capability of capturing spatial diversity patterns.
By contrast, our spatial predictions revealed drastic changes in the distribution of species composition, beta diversity and plant cover.
In effect, a large water surplus disrupted the strictly ordered vegetation assemblages which at first became apparent along the climatic gradient with wetter conditions.
Hence, spatial predictions might provide more detailed insights into ecosystem dynamics, especially in semi-arid ecosystems where changes in plant cover and distribution patterns are more pronounced than changes in alpha and gamma diversity.
>
>The influence of topographic and edaphic variables on plant diversity and productivity also gained in importance with wetter conditions but this effect was not further amplified with very humid conditions.
Our irrigation-fertilization experiment confirms these findings but also revealed the true beneficial effect of fertilizing under wet Super Niño conditions when plant cover was three times as high as in the unfertilized plots.
>
>In conclusion, our predictive mapping approach can identify locations suitable for restoration and conservation, and thus could form the basis for an informed conservation management (forest management plans).
At the same time, the results of our experiment can equip farmers with information regarding sustainable agrarian management which might increase the productivity on already existing farmland which in turn might help to prevent the destruction of the last remnants of TDF in the study area.

24. **Line 285: You write that grasses have a drought memory. So is it possible that the changes in biodiversity you observed are not due to current ENSO conditions but due to the conditions in the previous year(s)? This would change the outcome completely. Also, only one year per ENSO cycle was sampled. So I wonder if a general conclusion about the relationship between ENSO and biodiversity is justified.**

We are sorry for the confusion.
But in fact, we have largely focused on current-year precipitation effects on vegetation development.
The ordination and modeling results provide ample evidence for these effects.
Still, it might be possible that to a much lesser extent previous year precipitation patterns can also have an effect.
However, this was not the aim of the study, and therefore the collected data cannot provide evidence for such effects.
A former reviewer, your predecessor, acknowledged this already but still asked us to discuss possible time-lagged responses.
Therefore, we have now rewritten as follows:

> The ordination and modeling results of our study provide ample evidence how current-year precipitation affects biodiversity dynamics.
Still, it might be possible that to a much lesser extent previous year rainfall might also have an effect in the study area [@tenhumberg_timelagged_2018].
For instance, @dudney_lagging_2017 observed an increase in grass and a decrease in forb abundance when rainfall was high one year earlier.
Similarly, @walter_plants_2011 suggest that grasses might have a 'drought-memory'.

Since we have observed each ENSO episode only once, we cannot draw general conclusions, and we have not done so in the paper.
Instead we described the observed biodiversity patterns for each episode.
Despite that, it is relatively safe to assume that (though we do not claim so in the manuscript):

- the floristic composition remains uniform across the entire gradient during dry years.
- a certain order of vegetation formations appears along the gradient in more humid years.
- the strict order of the vegetation formations are partly interrupted in very humid years.

25. **326-327: Most important finding is that fertilizer increases biomass? Probably well known… By the way Figure 6 shows that plant coverage increases with water input more than when fertilized.  Where is the corresponding plot for biomass that was used for this statement? Also don't understand why talking about crops here when the study in general (though the experimental sites were placed somewhere else) was about vegetation plots where agricultural area was explicitly excluded (Line 89)**

Indeed, it is probably well known that fertilizing increases biomass.
However, the true impact of fertilizing in the study area only becomes apparent with quite humid conditions (Super Niño) when plant cover was three times higher than in the unfertilized plots.

Please note also that the corresponding figure is now in color (Fig. \@ref(fig:irrigation)).
Additionally, the solid salmon line and the dashed gray lines represent plant coverage in %.
By using the dashes, it is now visible that in the Control panel fertilization had no effect and that the two lines overlay each other.
Thanks for helping us to make the figure better understandable.

You are also right, that the figure actually shows the increase in plant cover which we now clarified in the text (see below).
Additionally, we now explicitly state in the discussion that agricultural activity increases and that our experiment might help to improve the productivity on the already existing farmland instead of making disappear the last TDF remnants found in the study area.
Please note that the experiment was carried out in the study area, please refer also to the reply to comment #12.
We have rewritten as follows:

>In fact, the water-nitrogen interaction can result in a plant coverage three times as high as achieved when adding the same amount of water but without additional nutrients (Fig. 6).
This is of utmost importance for sustainable yet productive agricultural management in the study area, where farming activity increases along the gradient.
Hence, instead of making disappear the last remnants of TDF in the study area, it would be desirable to increase the productivity on already existing farmland.
The decision on which crops (annual herbal and grass species) to grow and how much fertilizer to use could be further improved by reliable climate predictions on the occurrence of ENSO episodes.

```{r irrigation, echo=FALSE, out.width="100%", fig.align="center", warning=FALSE, message=FALSE, fig.cap="Results of the irrigation-fertilization experiment. The blue line refers to the water input, note that the left y-axis was log-transformed. The solid salmon and the gray dashed lines represent the observed plant coverage in \\% (right y-axis)."}
# include_graphics(here("figures/17_nutrient_irrigation_experiment.png"))
load(here("figures/17_nutrient_irrigation_experiment.rda"))

# to plot manually axes labels outside of the panel, we have to disable the
# clipping beforehand
my.setting = trellis.par.get("clip")
my.setting$panel = "off"
trellis.par.set("clip", my.setting)
# trellis.par.get("clip")

# plot
print(p_1)
```

26. **Section 4.4.4: This section seems to be off scope because none of your own results is coming in here.**

<!-- 
Reviewer#2 this time has been reviewer#1 in the previous round.
However, reviewer#2 in the first round asked us to discuss ENSO indices.
Hence, we can delete the corresponding section.
-->

To discuss especially findings supported by our data (as also requested by you earlier), we have deleted the corresponding section.

# Reply to reviewer #2

1. **I am glad that the authors have taken my (and of the reviewer 2) previous comments and criticisms seriously. I think the manuscript is greatly improved as a result. This paper is an interesting contribution to understand plant community ecology and their response to climate change using spatial and temporal data and the integration to macroecological and experimental approaches.**

Thank you very much.
We are glad to hear this.
And thank you again for taking the time to improve our manuscript with your valuable feedback.

2. **The introduction is focused on the effect of water availability gradient on floristic composition, but the authors analyzed simultaneously plant diversity and biomass (or cover). Then, I think that it is necessary to include another paragraph in the introduction to present the effects of climate and explicitly how water availability influences biomass productivity and dynamics (or development).**

We have rewritten the introduction as follows:

>> Naturally, water availability drives floristic composition and biomass production in the (semi-)arid regions such as NW Peru [@muenchow_coupling_2013; @muenchow_woody_2013].
In fact, water is probably the most important limiting abiotic factor in NW Peru's tropical dry forests where it controls plant recruitment, establishment and survival, seedling production, vegetation cover as well as species diversity [@balvanera_distribution_2011;@salazar_tree_2020].
However, water availability does not only vary in space but also in time, especially in regions heavily affected by the El Niño Southern Oscillation (ENSO).

3. **The authors also introduce the “ecosystem diversity” term, which is calculated as the axis values of a DCA analysis for each plot and subsequently, these are used to model floristic composition along a spatial gradient. Also, they analyzed “beta-diversity” as a range of DCA axes and compare them among years. Both approaches are based on the analysis of variation of the floristic composition obtained from a DCA and correspond to beta-diversity analyses (Anderson et al. 2011) but differ in the spatial scale in which are thought. On the one hand, the “beta-diversity” analyzed by the authors is the total variation on samples among plots and compare their range among years. Then, this approach is beta-diversity at the regional scale (i.e. the entire study area scale). On the other hand, DCA axes' values for each plot were used to modeling spatial variation of floristic composition. Here, the local variation of each plot was taken into account (i.e. plot scale). It is a difference among both approaches, and it may be clarified in the text.**

You are absolutely right.
To avoid misunderstandings, we now refrain from using the term ecosystem diversity, and now solely use floristic composition throughout the document.
We have rewritten the corresponding methods section as follows:

>One objective of this study was the spatio-temporal mapping of floristic composition represented by the scores of the first DCA axis along the climatic gradient in the study area.
Please note that the resulting prediction map not only shows species composition but also differences in beta diversity on the predicted pixel level.
This is because the units of a DCA axis are given in SDs (see also previous section).

4. **In this way, we recommend including an additional analysis to assess differences among years in the floristic composition. You can use a test of multivariate homogeneity of groups dispersions (see Anderson et al. 2006). Also, you can see Anderson et al. (2011) for a discussion among variation among samples and turnover in beta-diversity analyses.**

Thank you for this interesting suggestion.

In the method section we added following sentence:

>>To test if beta diversity and floristic composition, respectively, varied significantly between years, we analyzed the multivariate homogeneity of group dispersions [@anderson_multivariate_2006].

We added following sentence to the results:

>>Overall, there were no significant interannual differences in beta diversity in accordance with the test for multivariate homogeneity of groups dispersions (ANOVA *F*-test: 0.05069).

And we rewrote the discussion as follows...

>>Beta diversity (expressed as the range of the first DCA axis), increased with wetter conditions (Fig. 3 & Table 1.
However, beta diversity did no change significantly between years (section 3.1).
Concordantly, the change of the shape of the DCA point scatter between the relatively humid years (panels 2012 & 2016 of Fig. 3 and the very humid neutral year (panel 2017 of Fig. 3 can be most likely attributed to the increase of plant cover (see also next paragraph).

... and

>> Overall, our spatial mapping approach concedes detailed insights into how floristic composition and beta diversity distribution patterns change in response to varied rainfall patterns caused by different ENSO episodes.
Similarly, @lovelace_geocomputation_2019a (chapter 13) predicted floristic composition on a Peruvian fog oasis which revealed that the vegetation belts along the mountain slope are partly interrupted. 
Consequently, predictive mapping can contribute to a more nuanced understanding of ecosystem dynamics.
Naturally, such insights cannot be captured by global beta diversity measures such as the analysis of multivariate homogeneity of groups dispersions which in our case simply yielded not significant results (see sections 3.1 & 4.1).

Finally, we also added following sentences to the Conclusions:

>Beta diversity also increased with wetter conditions but did not change significantly between years in accordance with the test for multivariate homogeneity of groups dispersions.
As useful as such global measures are in general, they are in essence a single descriptive figure which naturally lack the capability of capturing spatial diversity patterns.
By contrast, our spatial predictions revealed drastic changes in the distribution of species composition and beta diversity.
In effect, a large water surplus disrupted the strictly ordered vegetation formations which at first became apparent along the climatic gradient with wetter conditions.
Hence, spatial predictions might provide more detailed insights into ecosystem dynamics, especially in semi-arid ecosystems where changes in plant cover and distribution patterns are more pronounced than changes in alpha and gamma diversity.

5. **Also, I recommend running a GAM between mean local alpha diversity and distance to the sea and even including a year as a categorical variable (and potentially their interaction) to add statistical support to the effect of spatial variation at this level of diversity.**

We have now also modeled alpha diversity as a function of the interaction between distance to the sea and year using a GAM and a spline smoother consuming three degrees of freedom (5-fold 100-repeated spatial cross-validation yielded an NRMSE of ~36%).
However, the corresponding prediction map becomes rather smooth with only a small amount of variation (Figure \@ref(fig:spri-dist)).
This is because distance to the sea remains the same independent of the year and simply because distance to the sea gradually increases from west to east.

```{r spri-dist, echo=FALSE, out.width="100%", fig.align="center", message=FALSE, warning=FALSE, fig.cap="Spatial prediction of species richness along the cross-section for different ENSO episodes using distance to the sea as predictor. The urban boundaries of the cities Paita, Piura and Chulucanas are displayed in black. The Piura River and its tributaries are represented by light blue lines with a black contour.",}
library("sp")
library("sf")
load(here("figures/15_spri_preds.rda"))
print(spri_dist)
```

Hence, we have again used NDVI and the interaction of NDVI and year as the predictors (Figure \@ref(fig:spri-ndvi)).
However, the spatial cross-validation is only partly convincing with an NRMSE of 40%.
Still, the predicted species patterns look reasonable.
We have put the spatial prediction map of alpha diversity into the Supplementary Information.
Please note that we additionally used topographic predictors (slope, catchment area, aspect, curvatures) to test if this would yield even better results when doing the spatial cross-validation.
However, this was not the case.

```{r spri-ndvi, echo=FALSE, out.width="100%", fig.align="center", message=FALSE, warning=FALSE, fig.cap="Spatial prediction of species richness along the cross-section for different ENSO episodes using the NDVI and year interaction as predictors (used model: GAM with a spline smoother with 3 degrees of freedom and a Poisson distribution; NRMSE of a 5-fold 100-repeated CV: 40\\%). The urban boundaries of the cities Paita, Piura and Chulucanas are displayed in black. The Piura River and its tributaries are represented by light blue lines with a black contour."}
print(spri_ndvi)
```

We have added the following sentence to the discussion:

>> Finally, the spatial prediction of species richness (Appendix 4) supports the findings of the floristic gradient mapping since it revealed similar spatial diversity patterns and dynamics.

6. **Finally, I think that the integration of macroecological and experimental approaches to this study is very interesting. Then, the results of local species richness of the experimental approach should be included in the main text to improve the discussion about the effect of water availability and nutrient enrichment on local diversity of plant communities.**

We have moved a revised version of the figure showing the development of species richness during the experiment from the Appendix to the main text (Fig. \@ref(fig:spri-exp)).
Additionally, we have added sentences to several sections throughout the document as follows:

Methods:

>> For presenting the results of our irrigation-fertilization experiment we simply used descriptive statistics by plotting the increase in the number of species, plant cover as well as water input against time for the duration of the experiment.

Results:

>> Species richness followed a similar pattern.
That means, highest numbers in species were recorded in the the fertilized Super-El Niño plots, though with just seven observed species the overall number remained low (Fig. 7).

Discussion:

>> Our irrigation-fertilization experiment confirms the findings of the landscape scale, i.e., the increase in plant cover was much more pronounced than the increase in species richness (see Figs. 6 & 7).

```{r spri-exp, echo=FALSE, fig.cap="Mean number of species per treatment level plotted for the duration of the experiment. Each dot represents a sampling visit during which among others the numbr of species was recorded.", fig.align="center"}
spri_exp = readRDS(here("images/17_irrigation_spri.rds"))
print(spri_exp)
```

7. **Line 16: “change and reshape” could be changed by modify**

We have deleted the corresponding sentence.

8. **Line 50-52: The prediction could be better explained.**

Please refer to the reply to comment #3 and also to the replies to comments #16-19 of reviewer #1.

9. **Line 86-87: The word “randomly” maybe removed because the sample design was stratified.**

Changed as requested.

10. **Line 111: change “determination” by identification**

Thanks for spotting.
Changed as requested.

11. **Line 130-133: The comparison of precipitation values among worldclim and local station could be interesting to visualize as supplementary information.**

In fact, we have not used the WorldClim data in our analysis which is why we have deleted the corresponding sentence from the manuscript.
Still, when comparing the climate station data with the WorldClim data it becomes apparent that WorldClim probably underestimates the precipitation west of Piura (middle blue dot in Figure \@ref(fig:wclim)).

```{r wclim, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="The black line represents the precipitation as derived from WorldClim along the transect. The points represent the annual median precipitation values at the climatic stations.", fig.width=6, fig.height=4, fig.align="center"}

# attach data
# env_data containing wordclim precip
env_data = readRDS(here("images/11_env_data.rds")) %>%
  tidyr::separate(col = "pnr", into = c("id", "year"), sep = "_")

# precipitation data of the climatic stations
con = dbConnect(RSQLite::SQLite(), here("data/tables.gpkg"))
# dbListTables(con)
precip = RSQLite::dbReadTable(con, "precipitation")
towns = st_read(con, "towns_jvs") 
dbDisconnect(con)

# get center of towns-polygon
piura = st_centroid(towns) %>%
  st_coordinates %>%
  as.data.frame %>%
  slice(2)

names(piura) = tolower(names(piura))
piura = cbind(piura, name = "piura")

xy = filter(env_data, id %in% c(1, 46) & year == 2011) %>% 
  dplyr::select(x, y)
xy = cbind(xy, "name" = c("paita", "chulu"))
xy = bind_rows(xy, piura) %>%
  arrange(x)

med_precip = group_by(precip, year, station) %>% 
  summarize(precip = sum(precip, na.rm = TRUE)) %>%
  group_by(station) %>% 
  summarize(precip = median(precip)) %>%
  inner_join(., xy, by = c("station" = "name")) %>% 
  arrange(precip)
med_precip$station = c("Paita", "Piura", "Chulucanas")

# ok, use coords of towns
ggplot(env_data) +
  geom_line(aes(x = x, y = prec)) +
  geom_point(data = med_precip, 
             aes(x = x, y = precip, color = station), cex = 1.5) +
  ylab(label = "Precipitation (mm)") +
  xlab(label = "Longitude") +
  theme_bw()
```

12. **Line 138-140: “This measure is given in units of standard deviation (SD) with four SDs corresponding to a complete species turnover”. I think that the number of SD that separate two samples with complete species turnover depends of the analyzed dataset and four SD is not a rule. Please add a citation for this sentence.**

We have clarified as follows:

>> Beta diversity is expressed as the range of the first Detrended Correspondence Analysis (DCA) axis (see below). 
The unit of a DCA axis is given in units of standard deviation (SD).
4 SDs correspond to a complete species turnover  while a difference between 1 to 1.4 SD units already correspond to half a change in species composition [@legendre_numerical_2012].

13. **Line 146: We should include species richness in your experimental approaches results**

Please refer to the reply to comment #6.

14. **Lines 195-196: The figures 2 and 3 could be switched to maintain the results consistency.**

Thanks for spotting.
Changed as requested.

15. **Lines 235: In the figure 5 should be included the year label of each panel.**

Thanks for spotting, they year labels should have been there in the first place.
Changed as requested.

16. **Lines 244: The results of species richness of the experimental approach may be included on the main text.**

Please refer to the reply to comments #6.

17. **Lines 262: Here, the author can include the name of dominant species to illustrate to readers into the ecological traits that increases which increases water availability in this study system. **

We have rewritten the sentence as follows:

>> Interestingly, only a few grass (e.g., *Aristida adscensionis*, *Eragrostis cilianensis*) and forb species (e.g., *Tiquilia paronychioides*) were in large parts responsible for this increase (see section 3.2).

# References