### 2020-08-10 text/ecography/round_3/title_page.Rmd

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: xelatex
    # float is needed for floating figures
    extra_dependencies: ["fontspec", "sectsty", "float"]
    toc: false
    fig_caption: yes
#    number_sections: true
header-includes:
  # adjust figure captions
  - \usepackage{caption}
  - \captionsetup{font={footnotesize,bf,singlespacing}}
  # Ecography requires double line spacing and line numbering
  - \usepackage{setspace}
  - \doublespacing
  #- \usepackage{lineno}
  #- \linenumbers

title: "Monitoring and predictive mapping of floristic biodiversity along a climatic gradient in ENSO's terrestrial core region, NW Peru"
#bibliography: transect_refs.bib
#csl: ecography.csl
# biblio-style: "natbib"
#link-citations: true
---

Jannes Muenchow^1^, Petra Dieker^2^, Thea Böttcher^3^, Jonas Brock^4^, Gregor Didenko^1^, Tobias Fremout^5^, Desiree Jakubka^6^, Anke Jentsch^7^, Daniel Nüst^8^, Michael Richter^9^, Eric Frank Rodríguez^10^, Rodolfo Arismendiz Rodríguez^11^, Rütger Rollenbeck^12^, Pablo Salazar Zarsosa^11^, Patrick Schratz^1^, Alexander Brenning^1^

^1^ Institute of Geography, Friedrich Schiller University Jena, Löbdergraben 32, 07743 Jena, Germany. jannes.muenchow\@uni-jena.de. ORCID: https://orcid.org/0000-0001-7834-4717

^2^ Thünen Institute, Institute of Biodiversity, Bundesallee 65, 38116 Braunschweig, Germany

^3^ Institute of Biology, Humboldt-University of Berlin, Philippstr. 13, 10115 Berlin, Germany

^4^ Department of Ecological Modelling^,^ Helmholtz Centre for Environmental Research -- UFZ, Permoserstr. 15, 04318 Leipzig, Germany

^5^ Division of Forest, Nature and Landscape, KU Leuven, Heverlee (Leuven), Belgium

^6^ Institute of Ecology and Evolution, Friedrich-Schiller University Jena, Dornburger Str. 159, 07743 Jena, Germany

^7^ Disturbance Ecology, BayCEER, University of Bayreuth, Universitätsstr. 30, 95440 Bayreuth, Germany

^8^ Institute for Geoinformatics, University of Münster, Heisenbergstraße 2
48149 Münster, Germany

^9^ Institute of Geography, Friedrich Alexander University Erlangen-Nuremberg, 91058 Wetterkreuz 15, Erlangen, Germany

^10^ Herbarium Truxillense (HUT), National University of Trujillo, Trujillo, Peru

^11^ Faculty of Engineering, University of Piura, Piura, Peru

^12^ Laboratory for Climatology and Remote Sensing (LCRS), Faculty of Geography, University of Marburg, Deutschhausstr.10, 35032 Marburg, Germany

\section*{Abstract}

The tropical dry forests of NW Peru are heavily shaped by the El Niño Southern Oscillation (ENSO), where especially El Niño brings rain to arid to semi-arid areas.
However, the resulting effects on biodiversity patterns remain largely unknown as well as the effect of environmental variables on the floristic composition under varying rainfall patterns.
Therefore, we studied the spatio-temporal effects of different ENSO episodes on floristic biodiversity along a climatic gradient ranging from the coastal desert to the Andean foothills.
We sampled 50 vegetation plots in four years representing different ENSO episodes.
To highlight the spatio-temporal changes in floristic composition and beta diversity across ENSO episodes, we predicted ordination scores with a Generalized Additive Model.
We applied variation partitioning to test if topographic or edaphic variables gained in importance during more humid ENSO episodes.
Additionally, we executed an irrigation-fertilization experiment to quantify the beneficial effects of the water-nutrient interaction under different simulated ENSO rainfall scenarios. 
Plant species richness increased under humid conditions during the humid La Niña (2012) and the moderate El Niño (2016), and slightly decreased under the very humid conditions during the coastal El Niño (2017).
The spatial prediction revealed that specific vegetation formations became more pronounced with increasing water input, but that a large water surplus led to the disruption of the strict order along the climatic gradient.
Edaphic and topographic variables gained in importance with increased water availability (2012 & 2016), however, this effect was not further amplified under very wet conditions (2017).
The experiment showed that plant cover under Super Niño conditions was three times higher when fertilized.
Overall, our spatial predictions concede detailed insights into spatio-temporal ecosystem dynamics in response to varying rainfall caused by different ENSO episodes while the results of the experiment can support farmers regarding a sustainable agrarian management.

<!--
Overall, our results can help to 1) protect local human populations from devastating ENSO effects through informed conservation management and the restoration of degraded ecosystems; 2) equip farmers with information regarding sustainable agrarian management.
-->

**Keywords**: tropical dry forest, Peru, ecosystem diversity, ordination, statistical learning, predictive mapping, irrigation-fertilization experiment, productivity

*Acknowledgments* - We thank Msc Luis Urbina Zapata and Msc Nestor Atarama for their excellent work in the field. 
We are grateful to NASA LP DAAC for granting access to remote sensing imagery (ASTER, MODIS).

*Funding* - We are also most thankful to the Deutsche Forschungsgemeinschaft (project Ri 370/19-1), a Friedrich Schiller University of Jena program to support junior researchers (project DRM/2017-05) and the project "Ecología Funcional de los Bosques Andaluces y Predicciones Sobre Sus Cambios Futuros (For-Change)" (UCO-27943) by Junta de Andalucía (Spain) and European FEDER funds for funding this work.

*Author contributions* - JM, MR and AJ conceived the ideas.
JM, TB, DJ, RAR and PSZ conducted the ﬁeld work and collected the data with the help of collaborators.
JM, PD, AB, TF, JB, GD, RR, DJ, PS, DN and EFR analyzed the data.
JM, DN and PS were responsible for establishing the research compendium.
JM led the writing with the assistance of all co-authors.